A thesis draft

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Add missing
Fix minor typos

After the deployment of Openstack and Openbaton, the next logical step is to try
and make the component talk to each other. First, we tried a simple \verb!iperf!
data transmission with two VNFs deployed by Openbaton as a common
hypervisor-based Openstack instances, as explained in~\cite{openbatonIperf}.
With this test completed, we started experimenting Openstack so as to use Docker
containers instead of traditional hypervisor instances. When we started
developing this solution, we adopted the Zun module~\cite{zunmoduleos}, that
aims to smoothly integrate Docker in Openstack. Unfortunately, we were not able
to make it work properly. In particular, it was not possible to open
communications ports on the Docker containers, hence making it impossible for
the VNFs to communicate. On top of that, we found difficulties regarding the
integration of Docker images in TOSCA definitions and Openstack. These two
issues are related: the TOSCA definition does not have any keyword to
specificate in any way that the image used for the VNF definition is a Docker
image, and Openstack, with the Zun plugin, at the time of the implementation did
not performed the image pull from any Docker registry. The impossibility to
easily set the containers ports and the lack of some support between the TOSCA
definition and the Openstack Zun plugin pushed us to drop Openstack as a
possible solution and to start looking for another valid VIM component.

@@

\subsection{Kubernetes}
Kubernetes, as already described in the technology section, is an orchestrator
created by Google able to create a software abstraction from the underlying
virtualization technology. This enables Kubernetes to create an integration with
Docker. Kubernetes has an architecture based on plug-ins, giving the possibility
to the users to change part of Kubernetes itself, such as the networking routing
strategy or the persistence layer. With this flexibility Kubernetes offered to
us what Docker Swarm was not able to give. Additionally, we developed a custom
Domain Name System (DNS) lookup process and distribute other internal components
in order to avoid requests bottleneck (as explained better in
Section~\ref{chap:archimple:sec:secondattempt:sub:scalability}).

@@

Kubernetes does not require a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
For the initial deployment we relied on an internal Openstack instance installed
at the University of Padua made of four nodes with 4GB of RAM, 10GB of SSD
storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
7). We performed the installation using the official Kubernetes repositories,
taking care of performing an accurate cluster configuration, as better explained
in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

@@

The impossibility to develop a custom VIM driver and VNFM manager, given the
resources of hand, led us to search for an alternative solution. Then, we
started to work directly on the VIM. The base approach was to create an API
Server which in future an Openbaton plugin could contact and simply deploy TOSCA
definitions, interacting with a set of RESTful APIs. We called this simple
backend \emph{Harbor}.

>>>>

After the deployment of Openstack and Openbaton, the next logical step is to try
and make the components talk to each other. First, we tried a simple
\verb!iperf! data transmission with two VNFs deployed by Openbaton as a common
hypervisor-based Openstack instances, as explained in~\cite{openbatonIperf}.
With this test completed, we started experimenting Openstack so as to use Docker
containers instead of traditional hypervisor instances. When we started
developing this solution, we adopted the Zun module~\cite{zunmoduleos}, that
aims to smoothly integrate Docker in Openstack. Unfortunately, we were not able
to make it work properly. In particular, it was not possible to open
communications ports on the Docker containers, hence making it impossible for
the VNFs to communicate. On top of that, we found difficulties regarding the
integration of Docker images in TOSCA definitions and Openstack. These two
issues are related: the TOSCA definition does not have any keyword to
specificate in any way that the image used for the VNF definition is a Docker
image, and Openstack, with the Zun plugin, at the time of the implementation did
not performed the image pull from any Docker registry. The impossibility to
easily set the containers ports and the lack of some support between the TOSCA
definition and the Openstack Zun plugin pushed us to drop Openstack as a
possible solution and to start looking for another valid VIM component.

@@

\subsection{Kubernetes}
Kubernetes, as already described in the technology section, is an orchestrator
created by Google able to create a software abstraction from the underlying
virtualization technology. This enables Kubernetes to create an integration with
Docker. Kubernetes has an architecture based on plug-ins, giving the possibility
to the users to change part of Kubernetes itself, such as the networking routing
strategy or the persistence layer. With this flexibility Kubernetes offered to
us what Docker Swarm was not able to give. Additionally, we developed a custom
Domain Name System (DNS) lookup process and distributed other internal
components in order to avoid requests bottleneck (as explained better in
Section~\ref{chap:archimple:sec:secondattempt:sub:scalability}).

@@

Kubernetes does not require a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
For the initial deployment we relied on an internal Openstack instance installed
at the University of Padua, creating four nodes with 4GB of RAM, 10GB of SSD
storage and 2 vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
7). We performed the installation using the official Kubernetes repositories,
taking care of performing an accurate cluster configuration complete with
additional modules (e.g. networking and distributed persistent storage), as
better explained in Appendix~\ref{kubernetes}. After the initial installation,
we obtained a configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

@@

The impossibility to develop a custom VIM driver and VNFM manager, given the
resources at hand, led us to search for an alternative solution. Then, we
started to work directly on the VIM. The base approach was to create an API
Server which in future an Openbaton plugin could contact and simply deploy TOSCA
definitions, interacting with a set of RESTful APIs. We called this simple
backend \emph{Harbor}.
